---
title: Week 6 worksheet
author: STA303
date: 22 Feb 2025
---


On the jupiter server we need to reinstall the packages from source, because the default versions aren't compatible
```{r packages}
install.packages("Matrix", type='source')
install.packages("TMB", type='source')
```

Today we'll consider the `MathAchieve` dataset from the `MEMSS` package

```{r theData}
install.packages("MEMSS")
data("MathAchieve", package='MEMSS')
head(MathAchieve)
```

A description of the data is in the package's help file

```
help("MathAchieve", package='MEMSS')
```

Notable information  from the help file is below

```
This data frame contains the following columns:

School an ordered factor identifying the school that the student                  
          attends                                                                      
                                                                                       
Minority a factor with levels ‘No’ ‘Yes’ indicating if the student                
          is a member of a minority racial group.                                      
                                                                                       
Sex a factor with levels ‘Male’ ‘Female’                                          
                                                                                       
SES a numeric vector of socio-economic status.                                    
                                                                                       
MathAch a numeric vector of mathematics achievement scores.                       
                                                                                       
MEANSES a numeric vector of the mean SES for the school.   
```     

Below is an intercept-only linear regression model

```{r aDumbModel}
res = glmmTMB::glmmTMB(MathAch ~ 1, data = MathAchieve)
knitr::kable(confint(res, full=TRUE), digits=2)
```

Let's look at a profile likelihood plot of the standard deviation parameter

```{r dumbProfile}
theProf = profile(res, parm='sigma')
plot(theProf[,c('.focal','value')])
Sprob = c(0.5, 0.8, 0.95, 0.99)
Squant = qchisq(Sprob, df=1)/2
abline(h=Squant, lty=3)
```

You'll notice the minimum of the profile likelihood is not at the MLE for `sigma` in the `confint` table.  This is because internally `glmmTMB` works with the log of the standard devition.   This can be checked by noticing the exponential of the `betadisp` parameter is the same as the estimate above.

```{r checkParams}
res$fit$par
sqrt(exp(res$fit$par['betad']))
```

We can make a profile likelihood plot of the SD as follows.

```{r profSd}
theProf$sd = sqrt(exp(theProf$.focal))
plot(theProf[,c('sd','value')])
```

Now fit a better model, using sex, each student's SES, and the school-average SES, as well as a school-level random effect.

```{r aGoodModel}

```

Now make a plot of the profile likelihood for the school-level standard deviation parameter.  The parameter name you need to pass to the `parm=` argument is the name listed in `rownames(vcov(res, full=TRUE))`, it's going to be something like `'theta|School'`.


```{r profileSd}

```

You'll find the minimum of the profile likelihood isn't the same as the MLE, check if exponentiating the `theta` parameter in `res$fit$par` gives you something sensible.

```{r checkParams2}

```

Make a profile likelihood plot with SD on the x axis.

```{r profileSdConvert}

```

Now compare the random effects model with a linear regression without the school-level random effect.  We shoud expect the confidence intervals for the `MEANSES` variable  to be much narrower if we don't include the random effect.  There are 7000 observations in the dataset but only 160 schools.  The `Sex` and `SES` variables vary within schools and their CI's won't change much when the random effect is added.  The `MEANSES` variable has one value per school, so the sample size is effectively 160 instead of 7000, so the CI's will be narrower.

```{r noRandomEffectsModel}

```

Use the `confint` function to compare the confidence intervals of the models with and without random effects.

```{r compareModels}

```


