---
title: Week 7 worksheet
author: STA303
date: 3 March 2025
---

On the jupyter server we need to reinstall the packages from source, because the default versions aren't compatible
```{r packages, eval=FALSE}
install.packages(c("Matrix","TMB", "glmmTMB"), type='source')
```

To use the `ranefPlot` functino you'll need the `Pmisc` package

```{r pmisc}
if(!requireNamespace("Pmisc", quietly=TRUE)) {
	install.packages("Pmisc", repos='http://r-forge.r-project.org')
}
```


Today we'll be looking at fuel efficiency of cars, using data from

`https://www.fueleconomy.gov/feg/download.shtml`

If Donald has shut the web site down you can see it at

`https://web.archive.org/web/20250222131930/https://www.fueleconomy.gov/feg/download.shtml`


Download and read in the data:
```{r carData}
cUrl =  'https://www.fueleconomy.gov/feg/epadata/vehicles.csv.zip'
# or https://web.archive.org/web/20250222131930/https://www.fueleconomy.gov/feg/epadata/vehicles.csv.zip
cFile = file.path(tempdir(), basename(cUrl))
download.file(cUrl, cFile)
cFile2 = unzip(cFile, exdir=tempdir())
x = read.table(cFile2, sep=',', header=TRUE, stringsAsFactors=FALSE)
```

A description of variables is at

`https://www.fueleconomy.gov/feg/ws/index.shtml#vehicle`

important variables are

- barrels08 - annual petroleum consumption in barrels 
- comb08 - combined Miles per Gallon 
- make - manufacturer (division)


Here's a plot of fuel consumption over time
```{r plotCarData}
plot(x$year, x$barrels08)
plot(factor(x$fuelType), x$barrels08)
```

There are a lot of electric and alternate fuel vehicles, we'll remove them.
```{r subset}
xSub = x[x$atvType =='' & x$fuelType != 'CNG', ]
plot(xSub$year, xSub$barrels08, log='y', pch=16, col='#00000003')
```

Recode some variables
```{r dataRecode}
# centre and scale time variable
xSub$decade = (xSub$year - 2000)/10
# cylinders, as factor
xSub$cylFac = relevel(factor(xSub$cylinders), '4')
# transmission type
xSub$transmission = factor(gsub(" .*", "", xSub$trany), 
	levels = c('Automatic','Manual'))
# type of vehicle (car, van, pickup, SUV)
xSub$type = gsub("^Standard | Type$|,|s$", "", 
	gsub("( - |/| )[24]wd", "", xSub$VClass, ignore.case=TRUE))
```

Consider the model

- $Y_{ij}$ is fuel consumption for car model $j$ from manufacturer $i$
- $X_{ij}$ is covariates (cylinders, time (in decades since 2000), and transmission type)
- $U_i$ is manufacturer $i$'s random effect, giving that manufacturer's fuel efficiency relative to similar cars with the same transmission and number of cylinders.

$$
\begin{aligned}
Y_{ij}|U \sim & \text{Gamma}(\mu_{ij}/\nu, \nu)\\
\mu_{ij} = & X_{ij} \beta + U_i\\
U \sim & \text{N}(0, \Sigma)
\end{aligned}
$$

Before we start, let's check the interpretation of the `sigma` parameter
```{r checkGamma}
mysim = data.frame(y = rgamma(1000, shape=25, rate = 1))
confint(glmmTMB::glmmTMB(y~1, data=mysim), full=TRUE)
```
on my laptop I get `sigma` close to 5, so `sigma` is the shape parameter and the inverse of the coefficient of variation.  


```{r gammaModel}
myFitMakeG = glmmTMB::glmmTMB(???)
```

plot the random effects using `Pmisc::ranefPlot`, use `maxNames=` to increase the number of car manufacturers shown.

```{r ranefPlotG}
Pmisc::ranefPlot(???)
```

What's the most fuel efficient car manufacturer?  Of the car companies you've heard of, what's most fuel efficient?  What's less efficient, a Rolls-Royce or a Hummer?  Use `glmmTMB::ranef`.

