---
title: Week 8 worksheet
author: STA303
date: 10 March 2025
---



To use the `seasonalFormula` function you'll need the most recent version of the `Pmisc` package (reinstall if your version is more than a week old)

```{r pmisc}
if(!requireNamespace("Pmisc", quietly=TRUE)) {
	install.packages("Pmisc", repos='http://r-forge.r-project.org')
}
```


The point of today's lab is to check if the trend in CO2 was affected by the following.
- The 1973 oil embargo (OPEC meeting 16 Oct 1973);
- The fall of the Berlin wall (8 Nov 1989), preceeding events which lead to the collapse of heavy industry in the eastern block countries;
- China joining the World Trade Organization (11 Dec 2001) which was followed by an explosive growth in China's GDP;
- The world financial crisis (bankrupcy of Lehman Brothers 15 Sept 2008); and
- The COVID-19 pandemic (WHO state of emergency 20 Jan 2000).


```{r importantDates}
importantDates = as.Date(c(
    oil = '1973/10/16',
    berlin = '1989/11/8',
    china = '2001/12/11',
    lehman = '2008/9/15',
    covid = '2020/1/20'
))
```


Data from
[scrippsco2.ucsd.edu/data/atmospheric_co2/mlo.html](https://scrippsco2.ucsd.edu/data/atmospheric_co2/mlo.html)


```{r co2Data, echo=FALSE}
cUrl = paste0('https://scrippsco2.ucsd.edu/assets/data/atmospheric/',
    'stations/flask_co2/daily/daily_flask_co2_mlo.csv')
cFile = basename(cUrl)
if(!file.exists(cFile)) download.file(cUrl, cFile)
co2s = read.table(cFile, header=TRUE, sep=',', skip=76, stringsAsFactors=FALSE,
  col.names=c('dateOrig','time','junk1','junk2', 'Nflasks','quality','co2'))
co2s$date = as.Date(co2s$dateOrig)
# remove low-quality measurements
co2s = co2s[co2s$quality==0,  ]
```

```{r plot}
plot(co2s[,c('date','co2')], log='y')
```

Fit an appropriate generalized additive model to the data

```{r fitGam}
co2s$dateInt = as.integer(co2s$date)
res = mgcv::gam()
```

create a data frame for forecasting

```{r newdata}
newdata = data.frame(
    date = seq(as.Date('1960/1/1'), as.Date('2030/1/1'), by = 'week')
)
```

make predictions of the trend (without the seasonal effect)

```{r gamPred}

```

plot the predicted smooth effect

```{r thePlot}

```


For a more thorough analysis of these date see
![Zhang, Brown and Stafford 2025](https://link.springer.com/article/10.1007/s11222-025-10565-2)
